Pinpointing a person's skin tone and translating it into a reliable foundation recommendation has long frustrated both consumers and the cosmetics industry. Conventional shade-matching leans on subjective visual judgment or narrow product ranges, yielding poor fits — especially for darker complexions. This project introduces \textbf{ShadeAI}, an end-to-end, AI-driven pipeline that automates skin-tone classification and foundation-shade suggestion through computer vision and perceptual color science.

\vspace{0.5cm}

The pipeline comprises six tightly sequenced stages: (1)~image preprocessing with automatic white-balance normalization, (2)~face localization via Haar Cascade classifiers, (3)~rule-based skin segmentation exploiting HSV and YCbCr color-space constraints, (4)~feature extraction — principally the Individual Typology Angle (ITA°) and CIELAB distribution statistics, (5)~classification against the ten-class Monk Skin Tone (MST) scale, and (6)~shade ranking powered by the CIEDE2000 perceptual-difference metric. Given a single facial photograph, ShadeAI returns the five closest foundation matches drawn from a curated catalogue of 187 shades spanning 14 major cosmetic brands.

\vspace{0.5cm}

Quantitative evaluation yields an average colour difference ($\Delta E$) of 1.361 — a figure 49.6\% below the commonly cited research target of $\Delta E < 2.7$ and 55\% tighter than the professional industry ceiling of $\Delta E < 3.0$. Across test subjects ranging from MST-03 to MST-09, the system delivers an 85.7\% excellent-match rate ($\Delta E < 3.0$) with a 100\% completion rate; 57.1\% of top-1 recommendations fall below the perceptual-indistinguishability threshold ($\Delta E < 1.0$). End-to-end inference completes in under one second on commodity CPU hardware, requiring neither training data nor deep-learning models — a trait that makes ShadeAI immediately deployable.

\vspace{0.5cm}

Three distinct access points serve different use cases: a command-line interface (CLI) optimized for batch runs, a Streamlit-based web front-end for guided interactive use, and a FastAPI REST endpoint for programmatic integration. Before any analysis begins, a built-in quality gate evaluates six image-health metrics — brightness, blur, colour cast, face presence, spatial resolution, and exposure distribution — flagging suboptimal inputs with actionable guidance. The underlying shade database deliberately over-indexes on deeper tones (MST-06 through MST-10), closing a gap that has historically plagued commercial shade finders.

\vspace{0.5cm}

Taken together, ShadeAI illustrates how classical vision algorithms, rigorous colorimetry, and thoughtful software design can converge on a socially meaningful product. By hitting research-grade accuracy in a zero-shot, CPU-only configuration, the system lowers the barrier to inclusive cosmetic recommendation. Its modular architecture leaves clear upgrade paths — deep-learning segmentation, expanded shade catalogues, live-camera capture, and augmented-reality try-on — for future iterations.

\vspace{0.5cm}

\textbf{Keywords:} Skin Tone Detection, Monk Skin Tone Scale, CIEDE2000, Foundation Recommendation, Computer Vision, Color Science, ITA°, CIELAB, Cosmetic AI, Inclusive Technology
