\chapter{Literature Review}

\section{Introduction to Skin Tone Detection}

Researchers in computer vision, clinical dermatology, and cosmetic science have pursued automated skin-tone detection for over three decades. Reliable classification feeds into applications as varied as UV-exposure risk assessment, adaptive camera processing, bias auditing in facial recognition, and — central to this work — personalised cosmetic recommendation. This chapter surveys the classification systems, color-science foundations, algorithmic techniques, and commercial products that define the current landscape, then pinpoints the gaps ShadeAI is designed to fill.

\section{Skin Tone Classification Systems}

\subsection{Fitzpatrick Skin Type Scale}

Dermatologist Thomas B.\ Fitzpatrick introduced his six-type scale in 1975 to predict a patient's sunburn-and-tanning response. The categories run from Type~I (very fair, always burns) through Type~VI (dark brown to black, never burns):

\begin{itemize}
    \item Type I: Very fair, always burns, never tans
    \item Type II: Fair, usually burns, tans minimally
    \item Type III: Medium, sometimes burns, tans uniformly
    \item Type IV: Olive, rarely burns, tans easily
    \item Type V: Brown, very rarely burns, tans very easily
    \item Type VI: Dark brown to black, never burns, tans very easily
\end{itemize}

For clinical UV work the scale remains serviceable, but its shortcomings in a cosmetic context are well documented. It was calibrated predominantly on Caucasian populations, compressing the enormous variety among darker skin tones into a single bin (Type~VI). Because it classifies by photoreactivity rather than visual appearance, two individuals who look markedly different may land in the same Fitzpatrick type — rendering the scale nearly useless for shade matching.

\subsection{Monk Skin Tone Scale}

Dr.\ Ellis Monk at Harvard University developed the Monk Skin Tone (MST) scale specifically to address these representation failures. Published in 2019 and subsequently adopted by Google for fairness benchmarking, MST expands coverage to ten classes:

\begin{table}[h]
\centering
\begin{tabular}{|c|l|l|}
\hline
\textbf{Class} & \textbf{Label} & \textbf{Description} \\
\hline
MST-01 & Porcelain & Very fair, cool/pink undertones \\
MST-02 & Fair & Fair, warm/peachy undertones \\
MST-03 & Light & Light beige \\
MST-04 & Light-Medium & Light medium \\
MST-05 & Medium & Medium beige \\
MST-06 & Medium-Dark & Warm medium \\
MST-07 & Tan & Tan/olive \\
MST-08 & Brown & Brown \\
MST-09 & Deep Brown & Deep brown \\
MST-10 & Deep & Deep/ebony \\
\hline
\end{tabular}
\caption{Monk Skin Tone Scale Classification}
\end{table}

Four of those ten bins (MST-07 through MST-10) are dedicated to brown and deep complexions — a four-fold increase in granularity over Fitzpatrick's single dark category. That finer resolution maps naturally onto cosmetic product lines, where adjacent shades may differ by only a few $\Delta E$ units.

\subsection{Individual Typology Angle (ITA°)}

Chardon et al.\ (1991) proposed ITA° as a continuous, device-independent metric for skin lightness. It is computed from CIELAB coordinates:

\begin{equation}
ITA° = \arctan\left(\frac{L^* - 50}{b^*}\right) \times \frac{180}{\pi}
\end{equation}

where $L^*$ captures lightness and $b^*$ captures the yellow–blue axis. High ITA° values indicate very fair skin; strongly negative values correspond to the deepest tones. Because ITA° is a scalar derived from perceptually grounded coordinates, it bridges the gap between continuous colour measurement and discrete classification — a threshold lookup can map it directly onto MST classes.

\section{Color Science and Color Spaces}

\subsection{RGB Color Space}

Digital cameras natively produce RGB (Red, Green, Blue) triplets. The space is intuitive and ubiquitous, yet it suffers from two fundamental drawbacks for skin analysis. First, it is device-dependent: the same scene yields different RGB values on different sensors. Second, it is perceptually non-uniform — a ten-unit shift in the red channel may be imperceptible in one region of the gamut yet glaringly obvious in another. These properties disqualify raw RGB from any reliable colour-matching pipeline.

\subsection{CIELAB Color Space}

The Commission Internationale de l'Éclairage defined CIELAB in 1976 to remedy exactly those problems. Three orthogonal axes encode colour:

\begin{itemize}
    \item $L^*$: Lightness (0 = black, 100 = white)
    \item $a^*$: Red-green axis (negative = green, positive = red)
    \item $b^*$: Yellow-blue axis (negative = blue, positive = yellow)
\end{itemize}

CIELAB suits skin-tone work for several reasons:
\begin{enumerate}
    \item Lightness and chromaticity are separable, so tone and undertone can be analyzed independently
    \item Distances in the space approximate perceived differences — a property called perceptual uniformity
    \item Measurements are device-independent once the illuminant white-point is fixed
    \item The space underpins industry-standard colour-difference formulas, including CIEDE2000
\end{enumerate}

\subsection{HSV and YCbCr Color Spaces}

Empirical studies have long noted that skin pixels cluster tightly in HSV (Hue, Saturation, Value) and YCbCr (Luma, Blue-difference, Red-difference) spaces, far more compactly than in RGB. Typical skin regions satisfy:

\begin{itemize}
    \item HSV: Skin typically has hue values between 0-50° and saturation between 0.2-0.6
    \item YCbCr: Skin pixels satisfy conditions like $77 \leq Cb \leq 127$ and $133 \leq Cr \leq 173$
\end{itemize}

Both representations decouple luminance from chrominance, which makes the thresholds robust across a wide range of indoor and outdoor lighting conditions — a decisive advantage for rule-based segmentation.

\subsection{Color Difference Metrics}

\subsubsection{CIEDE2000}

CIEDE2000 superseded earlier CIE delta-E formulas ($\Delta E^*_{ab}$, $\Delta E^*_{94}$) in the year 2000. It incorporates several perceptual corrections absent from its predecessors:

\begin{itemize}
    \item Weighting functions for lightness, chroma, and hue
    \item Rotation term for blue region
    \item Compensation for neutral colors
    \item Perceptual uniformity corrections
\end{itemize}

The formula is algebraically involved, but the practical interpretation is straightforward. A $\Delta E$ below 1.0 is invisible to a trained observer; below 2.0 qualifies as outstanding; below 3.0 is the accepted threshold for professional cosmetic matching. Published shade-matching research typically targets $\Delta E < 3.0$, and commercial tools rarely quantify their error at all.

\section{Computer Vision Techniques}

\subsection{Face Detection}

Isolating the face is the indispensable first step — without it, background pixels, clothing, and hair contaminate the colour statistics.

\subsubsection{Haar Cascade Classifiers}

Viola and Jones (2001) popularized Haar-feature cascades, which remain among the fastest face detectors available:
\begin{itemize}
    \item Real-time throughput on a single CPU core
    \item Tolerance to moderate pose and lighting variation
    \item Pre-trained models bundled with OpenCV
    \item Strong accuracy on frontal and near-frontal faces
\end{itemize}

The main weaknesses are degraded recall on profile views and sensitivity to heavy occlusion (sunglasses, scarves).

\subsubsection{Multi-task Cascaded Convolutional Networks (MTCNN)}

Zhang et al.\ (2016) introduced MTCNN, a three-stage deep detector that jointly localises faces and predicts five facial landmarks:
\begin{itemize}
    \item Substantially higher recall than Haar cascades on challenging poses
    \item Landmark output enables alignment and region-of-interest refinement
    \item Robust to expression, partial occlusion, and scale variation
    \item Confidence scores accompany every detection
\end{itemize}

The trade-off is cost: MTCNN demands considerably more compute and may need GPU acceleration to hit real-time frame rates.

\subsection{Skin Segmentation}

Once the face bounding box is known, the next task is to separate skin pixels from non-skin regions — eyes, brows, lips, and any visible background.

\subsubsection{Rule-Based Methods}

Threshold-driven approaches exploit the tight chromatic clustering of skin:

\begin{itemize}
    \item HSV thresholding: $0° \leq H \leq 50°$, $0.2 \leq S \leq 0.6$
    \item YCbCr thresholding: $77 \leq Cb \leq 127$, $133 \leq Cr \leq 173$
    \item Combined methods using multiple color spaces
\end{itemize}

The appeal is obvious: zero training, minimal latency, and no data dependency. The flip side is brittleness under extreme lighting and reduced precision across the full skin-tone gamut.

\subsubsection{Machine Learning Methods}

Learned classifiers model the skin–non-skin boundary from labelled pixel samples:

\begin{itemize}
    \item Gaussian Mixture Models (GMM)
    \item Support Vector Machines (SVM)
    \item Random Forests
    \item Neural Networks
\end{itemize}

Accuracy climbs relative to hand-crafted thresholds, but the models require curated training sets and incur higher inference cost.

\subsubsection{Deep Learning Methods}

Fully convolutional architectures set the current accuracy bar for pixel-level skin parsing:

\begin{itemize}
    \item U-Net: Encoder-decoder architecture with skip connections
    \item SegNet: Similar architecture with pooling indices
    \item DeepLab: Uses atrous convolution for multi-scale features
\end{itemize}

These networks achieve the tightest masks, yet they demand large annotated corpora, GPU training, and longer iteration cycles — resources that put them out of reach for lightweight or rapid-deployment scenarios.

\section{Existing Skin Tone Detection Systems}

\subsection{Academic Research}

\subsubsection{Deep Learning Approaches}

Several recent studies lean heavily on convolutional architectures:

\begin{itemize}
    \item \textbf{EfficientNet-based Classification:} Fine-tuned EfficientNet variants report 92–94\% top-1 accuracy on MST-labelled test sets, but they need thousands of annotated face crops and GPU-hours for training.
    
    \item \textbf{CNN + SVM Hybrid:} A two-stage pipeline extracts CNN features then feeds them to an SVM head. Published $\Delta E$ values hover around 3.0–3.5 for foundation matching — acceptable but not impressive.
    
    \item \textbf{ResNet-based Systems:} ResNet-50 backbones have achieved 88–90\% accuracy on the six-class Fitzpatrick scale, though the coarser taxonomy inflates reported accuracy.
\end{itemize}

Across the board, shared weaknesses emerge:
\begin{itemize}
    \item Steep computational demands
    \item Hunger for large, expertly annotated datasets
    \item Shade databases rarely exceeding 50–80 entries
    \item No published deployable front-end or API
\end{itemize}

\subsubsection{Traditional Computer Vision Approaches}

Classical pipelines have also been explored, typically combining:

\begin{itemize}
    \item ITA°-based classification achieving 85-90\% accuracy
    \item Rule-based segmentation with morphological operations
    \item Color histogram analysis for undertone detection
\end{itemize}

They are faster and training-free, but published results show accuracy dips on heavily shadowed or very dark-toned images.

\subsection{Commercial Systems}

\subsubsection{L'Oréal Shade Finder}

L'Oréal's mobile app analyses a selfie and maps it to the brand's product catalogue:
\begin{itemize}
    \item Real-time camera analysis
    \item Integration with product catalog
    \item Virtual try-on capabilities
\end{itemize}

No public documentation describes the model architecture or training data; the technology is entirely proprietary.

\subsubsection{Sephora Virtual Artist}

Sephora's AR-based tool focuses on visual engagement:
\begin{itemize}
    \item Augmented reality overlay
    \item Multiple product categories
    \item Integration with e-commerce
\end{itemize}

Anecdotal user reports suggest the emphasis is on visual appeal rather than colorimetric precision.

\subsubsection{Perfect Corp. YouCam Makeup}

YouCam Makeup bundles a broad feature set:
\begin{itemize}
    \item Real-time face tracking
    \item Multiple makeup products
    \item Skin analysis features
\end{itemize}

Like its competitors, the system relies on undisclosed proprietary models and does not publish quantitative $\Delta E$ benchmarks.

\subsection{Limitations of Existing Systems}

Synthesising the academic and commercial landscape reveals a recurring set of shortcomings:

\begin{enumerate}
    \item \textbf{Accuracy:} Reported $\Delta E$ values cluster between 3.0 and 5.0 — passable but far from imperceptible. Deeper tones consistently score worse.
    
    \item \textbf{Accessibility:} Deep-learning pipelines presuppose GPU hardware and ML expertise, locking out solo developers and small-scale beauty brands.
    
    \item \textbf{Training Data:} Supervised methods depend on thousands of labelled images — expensive to collect and fraught with annotation inconsistency.
    
    \item \textbf{Database Coverage:} Most shade databases cap at 50–100 entries and under-represent MST-07 through MST-10.
    
    \item \textbf{Production Readiness:} Academic codebases rarely ship with a user-facing interface, deployment guide, or automated quality gate.
    
    \item \textbf{Inclusivity:} A surprising number of systems still default to the six-bin Fitzpatrick taxonomy, forfeiting the granularity needed for accurate deep-tone matching.
\end{enumerate}

\section{Research Gap and Opportunity}

\subsection{Gaps in Current Research}

\begin{enumerate}
    \item \textbf{Accessibility Gap:} High-accuracy systems overwhelmingly require GPU hardware and large-scale training, shutting out resource-constrained developers.
    
    \item \textbf{Database Gap:} Existing shade catalogues are thin and skewed light. Deeper MST classes are chronically under-represented.
    
    \item \textbf{Implementation Gap:} The distance from a notebook prototype to a production endpoint remains vast; almost no published work closes it.
    
    \item \textbf{Inclusivity Gap:} Fitzpatrick persists as the default classification backbone despite well-known coverage deficiencies.
    
    \item \textbf{Documentation Gap:} Reproducibility suffers because papers seldom release comprehensive code, data, or configuration details.
\end{enumerate}

\subsection{Opportunities for Innovation}

\begin{enumerate}
    \item \textbf{Classical Methods:} A carefully engineered pipeline of white-balance correction, dual-space segmentation, ITA° classification, and CIEDE2000 matching can rival deep-learning accuracy — without the infrastructure overhead.
    
    \item \textbf{Comprehensive Database:} Curating 180+ shades with lab-grade CIELAB coordinates, undertone tags, and MST-range metadata directly tightens $\Delta E$.
    
    \item \textbf{Inclusive Classification:} Anchoring on the Monk scale aligns the system with the industry's broader push toward equitable representation.
    
    \item \textbf{Production-Ready System:} Shipping CLI, web, and API interfaces alongside full documentation bridges the prototype-to-product chasm.
    
    \item \textbf{Open Source:} Releasing the system publicly invites community scrutiny, contributions, and downstream adoption.
\end{enumerate}

\section{Theoretical Foundation}

\subsection{Color Perception and Matching}

Human colour perception is shaped by several interacting factors:

\begin{itemize}
    \item \textbf{Lighting:} The same object appears different under different illuminants (metamerism)
    \item \textbf{Surrounding Colors:} Adjacent colors affect perception (simultaneous contrast)
    \item \textbf{Individual Variation:} Color perception varies between individuals
    \item \textbf{Context:} Expectations and context influence color judgment
\end{itemize}

Foundation matching aims to minimise the perceived gap between product and skin under typical viewing illuminants — daylight and warm indoor light. CIEDE2000 is the best available mathematical proxy for that perceptual gap, which is why it serves as the ranking metric in ShadeAI.

\subsection{White Balance and Color Constancy}

Camera sensors record the scene illuminant alongside the subject, injecting a colour cast that can shift extracted skin coordinates by several CIELAB units. White-balance algorithms attempt to subtract this cast. The Gray World assumption — that the spatial average of a natural scene should be achromatic — is one of the oldest and simplest normalisations. It requires no calibration target, no training, and runs in constant time per pixel. For the controlled-lighting conditions typical of selfies, it proves surprisingly effective.

\subsection{Skin Tone and Undertone}

Tone describes how light or dark the skin appears; undertone captures the hue that shows through:

\begin{itemize}
    \item \textbf{Warm Undertone:} Yellow, golden, or peachy hues (high $b^*$ in CIELAB)
    \item \textbf{Cool Undertone:} Pink, red, or bluish hues (low $b^*$, high $a^*$)
    \item \textbf{Neutral Undertone:} Balanced mix of warm and cool
\end{itemize}

Getting the tone right but the undertone wrong produces the telltale "mask effect" — the foundation sits visibly on top of the skin rather than blending in. Any effective matching system must account for both dimensions.

\section{Summary}

This chapter has surveyed classification systems, colour-science fundamentals, algorithmic approaches, and existing products to establish the terrain ShadeAI occupies. Five key takeaways anchor the design decisions in the chapters that follow:

\begin{enumerate}
    \item The Monk Skin Tone scale offers meaningfully finer granularity than the Fitzpatrick scale, particularly for darker complexions
    \item CIELAB coordinates and the CIEDE2000 metric furnish the perceptual backbone needed for sub-3.0 $\Delta E$ matching
    \item Both classical and deep-learning pipelines have demonstrated viability, each with clear trade-offs in accuracy, cost, and deployment readiness
    \item Existing systems — academic and commercial alike — leave significant gaps in accessibility, database breadth, and production maturity
    \item A classical pipeline paired with a comprehensive shade database can, in principle, close those gaps without demanding GPU resources or labelled training data
\end{enumerate}

The next chapter translates these observations into a concrete system architecture and algorithmic methodology.
