\newpage
\thispagestyle{empty}
\begin{center}
\textbf{\Large DECLARATION}
\end{center}

\vspace{1cm}

We hereby declare that the project work entitled \textbf{"ShadeAI: AI-Based Skin Tone and Undertone Detection for Cosmetic Shade Recommendation"} submitted to JSS Academy of Technical Education, Noida, in partial fulfillment of the requirements for the award of the degree of Bachelor of Technology in Computer Science and Engineering (AI \& ML) is an authentic record of our own work carried out during the period from August 2025 to April 2026 under the supervision of \textbf{Ms. Vishakha Chauhan}, Department of Computer Science and Engineering-AIML, JSS Academy of Technical Education, Noida.

\vspace{0.5cm}

The matter embodied in this project report has not been submitted by us for the award of any other degree or diploma.

\vspace{2cm}

\begin{flushleft}
\textbf{Khushi Yadav} \hspace{5cm} \textbf{Chinmay Nanda}\\
Roll No: 22CSAIML0XX \hspace{5.5cm} Roll No: 22CSAIML076\\
\vspace{1cm}
\textbf{Ayush Raj Chauhan}\\
Roll No: 22CSAIML0XX\\
\end{flushleft}

\vspace{1cm}

\textbf{Date:} May 2026\\
\textbf{Place:} Noida

\newpage

\thispagestyle{empty}
\begin{center}
\textbf{\Large CERTIFICATE}
\end{center}

\vspace{1cm}

This is to certify that the project work entitled \textbf{"ShadeAI: AI-Based Skin Tone and Undertone Detection for Cosmetic Shade Recommendation"} is a bonafide record of work carried out by \textbf{Khushi Yadav (Roll No. 22CSAIML0XX), Chinmay Nanda (Roll No. 22CSAIML076), and Ayush Raj Chauhan (Roll No. 22CSAIML0XX)} in partial fulfillment of the requirements for the award of the degree of Bachelor of Technology in Computer Science and Engineering (AI \& ML) from Dr. APJ Abdul Kalam Technical University, Lucknow, during the academic year 2025-2026.

\vspace{2cm}

\begin{flushleft}
\textbf{Ms. Vishakha Chauhan} \hspace{5cm} \textbf{Dr. [HOD Name]}\\
Project Supervisor \hspace{6cm} Head of Department\\
Department of CSE-AIML \hspace{4.5cm} Department of CSE-AIML\\
JSS Academy of Technical Education \hspace{2.5cm} JSS Academy of Technical Education\\
\end{flushleft}

\vspace{2cm}

\textbf{External Examiner}

\vspace{1cm}

\textbf{Date:} \_\_\_\_\_\_\_\_\_\_\_\_\_\_

\newpage

\thispagestyle{empty}
\begin{center}
\textbf{\Large ACKNOWLEDGEMENT}
\end{center}

\vspace{1cm}

We would like to express our sincere gratitude to all those who have contributed to the successful completion of this project.

\vspace{0.5cm}

First and foremost, we are deeply grateful to our project supervisor, \textbf{Ms. Vishakha Chauhan}, for their invaluable guidance, continuous support, and encouragement throughout the project. Their expertise and insights have been instrumental in shaping this work.

\vspace{0.5cm}

We extend our heartfelt thanks to \textbf{Dr. [HOD Name]}, Head of the Department of Computer Science and Engineering-AIML, for providing us with the necessary facilities and resources to carry out this project.

\vspace{0.5cm}

We are thankful to \textbf{Dr. [Director Name]}, Director of JSS Academy of Technical Education, Noida, for creating an environment conducive to learning and research.

\vspace{0.5cm}

We would like to acknowledge the faculty members of the Department of CSE-AIML for their valuable suggestions and support during various stages of the project.

\vspace{0.5cm}

We are grateful to our parents and family members for their constant encouragement and moral support throughout our academic journey.

\vspace{0.5cm}

We also thank our friends and colleagues who have directly or indirectly contributed to the completion of this project.

\vspace{0.5cm}

Finally, we acknowledge the researchers and developers whose published work and open-source contributions have been invaluable references for this project.

\vspace{2cm}

\begin{flushright}
\textbf{Khushi Yadav}\\
\textbf{Chinmay Nanda}\\
\textbf{Ayush Raj Chauhan}\\
\end{flushright}

\newpage
