\section*{References}

\begin{enumerate}

\item Chardon, A., Cretois, I., \& Hourseau, C. (1991). Skin colour typology and suntanning pathways. \textit{International Journal of Cosmetic Science}, 13(4), 191-208.

\item Monk, E. P. (2019). Monk Skin Tone Scale: A more inclusive representation of skin tones. \textit{Harvard University}.

\item Commission Internationale de l'Éclairage (CIE). (2001). Improvement to industrial colour-difference evaluation. \textit{CIE Technical Report 142-2001}.

\item Luo, M. R., Cui, G., \& Rigg, B. (2001). The development of the CIE 2000 colour-difference formula: CIEDE2000. \textit{Color Research \& Application}, 26(5), 340-350.

\item Viola, P., \& Jones, M. (2001). Rapid object detection using a boosted cascade of simple features. \textit{Proceedings of the 2001 IEEE Computer Society Conference on Computer Vision and Pattern Recognition}, 1, I-I.

\item Zhang, K., Zhang, Z., Li, Z., \& Qiao, Y. (2016). Joint face detection and alignment using multitask cascaded convolutional networks. \textit{IEEE Signal Processing Letters}, 23(10), 1499-1503.

\item Ronneberger, O., Fischer, P., \& Brox, T. (2015). U-Net: Convolutional networks for biomedical image segmentation. \textit{International Conference on Medical Image Computing and Computer-Assisted Intervention}, 234-241.

\item Tan, M., \& Le, Q. (2019). EfficientNet: Rethinking model scaling for convolutional neural networks. \textit{International Conference on Machine Learning}, 6105-6114.

\item Bradski, G. (2000). The OpenCV Library. \textit{Dr. Dobb's Journal of Software Tools}.

\item Pedregosa, F., et al. (2011). Scikit-learn: Machine learning in Python. \textit{Journal of Machine Learning Research}, 12, 2825-2830.

\item Ramírez, S. (2018). FastAPI framework, high performance, easy to learn, fast to code, ready for production. \textit{https://fastapi.tiangolo.com/}

\item Streamlit Inc. (2019). Streamlit: The fastest way to build data apps. \textit{https://streamlit.io/}

\item Fitzpatrick, T. B. (1975). Soleil et peau. \textit{Journal de Médecine Esthétique}, 2, 33-34.

\item Kolkur, S., Kalbande, D., Shimpi, P., Bapat, C., \& Jatakia, J. (2017). Human skin detection using RGB, HSV and YCbCr color models. \textit{International Conference on Communication and Signal Processing}, 324-332.

\item Kakumanu, P., Makrogiannis, S., \& Bourbakis, N. (2007). A survey of skin-color modeling and detection methods. \textit{Pattern Recognition}, 40(3), 1106-1122.

\item Del Bino, S., \& Bernerd, F. (2013). Variations in skin colour and the biological consequences of ultraviolet radiation exposure. \textit{British Journal of Dermatology}, 169(s3), 33-40.

\item Jablonski, N. G., \& Chaplin, G. (2000). The evolution of human skin coloration. \textit{Journal of Human Evolution}, 39(1), 57-106.

\item Baranoski, G. V., \& Krishnaswamy, A. (2010). \textit{Light \& Skin Interactions: Simulations for Computer Graphics Applications}. Morgan Kaufmann.

\item Fairchild, M. D. (2013). \textit{Color Appearance Models} (3rd ed.). John Wiley \& Sons.

\item Hunt, R. W. G., \& Pointer, M. R. (2011). \textit{Measuring Colour} (4th ed.). John Wiley \& Sons.

\item Sharma, G., Wu, W., \& Dalal, E. N. (2005). The CIEDE2000 color-difference formula: Implementation notes, supplementary test data, and mathematical observations. \textit{Color Research \& Application}, 30(1), 21-30.

\item Angelopoulou, E., Molana, R., \& Daniilidis, K. (2001). Multispectral skin color modeling. \textit{Proceedings IEEE Conference on Computer Vision and Pattern Recognition}, 2, II-635.

\item Phung, S. L., Bouzerdoum, A., \& Chai, D. (2005). Skin segmentation using color pixel classification: Analysis and comparison. \textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}, 27(1), 148-154.

\item Nishidate, I., Aizu, Y., \& Mishina, H. (2004). Estimation of melanin and hemoglobin in skin tissue using multiple regression analysis aided by Monte Carlo simulation. \textit{Journal of Biomedical Optics}, 9(4), 700-710.

\item Tsumura, N., Ojima, N., Sato, K., Shiraishi, M., Shimizu, H., Nabeshima, H., ... \& Miyake, Y. (2003). Image-based skin color and texture analysis/synthesis by extracting hemoglobin and melanin information in the skin. \textit{ACM Transactions on Graphics}, 22(3), 770-779.

\item Google AI. (2022). Monk Skin Tone Scale: A more inclusive approach to skin tone research. \textit{Google AI Blog}.

\item Buolamwini, J., \& Gebru, T. (2018). Gender shades: Intersectional accuracy disparities in commercial gender classification. \textit{Conference on Fairness, Accountability and Transparency}, 77-91.

\item Raji, I. D., \& Buolamwini, J. (2019). Actionable auditing: Investigating the impact of publicly naming biased performance results of commercial AI products. \textit{Proceedings of the 2019 AAAI/ACM Conference on AI, Ethics, and Society}, 429-435.

\item Kärkkäinen, K., \& Joo, J. (2021). FairFace: Face attribute dataset for balanced race, gender, and age for bias measurement and mitigation. \textit{Proceedings of the IEEE/CVF Winter Conference on Applications of Computer Vision}, 1548-1558.

\item Merler, M., Ratha, N., Feris, R. S., \& Smith, J. R. (2019). Diversity in faces. \textit{arXiv preprint arXiv:1901.10436}.

\item Deng, J., Guo, J., Xue, N., \& Zafeiriou, S. (2019). ArcFace: Additive angular margin loss for deep face recognition. \textit{Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition}, 4690-4699.

\end{enumerate}

\subsection*{Online Resources}

\begin{enumerate}[resume]

\item OpenCV Documentation. (2024). Face Detection using Haar Cascades. Retrieved from https://docs.opencv.org/

\item scikit-image Documentation. (2024). Color space conversions. Retrieved from https://scikit-image.org/

\item NumPy Documentation. (2024). Array programming with NumPy. Retrieved from https://numpy.org/

\item FastAPI Documentation. (2024). FastAPI framework documentation. Retrieved from https://fastapi.tiangolo.com/

\item Streamlit Documentation. (2024). Streamlit documentation. Retrieved from https://docs.streamlit.io/

\item Python Software Foundation. (2024). Python 3.10 Documentation. Retrieved from https://docs.python.org/3.10/

\item Docker Documentation. (2024). Docker containerization platform. Retrieved from https://docs.docker.com/

\item Git Documentation. (2024). Git version control system. Retrieved from https://git-scm.com/doc

\item GitHub. (2024). GitHub code hosting platform. Retrieved from https://github.com/

\item Stack Overflow. (2024). Developer community and Q\&A. Retrieved from https://stackoverflow.com/

\end{enumerate}

\subsection*{Cosmetic Brand Resources}

\begin{enumerate}[resume]

\item MAC Cosmetics. (2024). Studio Fix Fluid Foundation shade range. Retrieved from https://www.maccosmetics.com/

\item Fenty Beauty. (2024). Pro Filt'r Soft Matte Longwear Foundation. Retrieved from https://fentybeauty.com/

\item NARS Cosmetics. (2024). Sheer Glow Foundation. Retrieved from https://www.narscosmetics.com/

\item Maybelline. (2024). Fit Me Matte + Poreless Foundation. Retrieved from https://www.maybelline.com/

\item L'Oréal Paris. (2024). True Match Foundation. Retrieved from https://www.lorealparisusa.com/

\item Estée Lauder. (2024). Double Wear Stay-in-Place Foundation. Retrieved from https://www.esteelauder.com/

\item Bobbi Brown. (2024). Skin Long-Wear Weightless Foundation. Retrieved from https://www.bobbibrowncosmetics.com/

\item Lancôme. (2024). Teint Idole Ultra Wear Foundation. Retrieved from https://www.lancome-usa.com/

\item Black Opal. (2024). True Color Foundation. Retrieved from https://blackopalbeauty.com/

\item Iman Cosmetics. (2024). Second to None Cream to Powder Foundation. Retrieved from https://imancosmetics.com/

\end{enumerate}

\subsection*{Standards and Specifications}

\begin{enumerate}[resume]

\item International Commission on Illumination (CIE). (1976). Colorimetry. \textit{CIE Publication 15}.

\item International Commission on Illumination (CIE). (1986). Colorimetry (2nd ed.). \textit{CIE Publication 15.2}.

\item International Organization for Standardization. (1999). Graphic technology — Process control for the production of half-tone colour separations, proof and production prints — Part 1: Parameters and measurement methods. \textit{ISO 12647-1:1996}.

\item sRGB Color Space Specification. (1996). IEC 61966-2-1:1999.

\item World Wide Web Consortium (W3C). (2011). CSS Color Module Level 3. \textit{W3C Recommendation}.

\end{enumerate}
