\newpage
\thispagestyle{empty}
\begin{center}
\textbf{\Large VISION AND MISSION}
\end{center}

\vspace{1cm}

\section*{Vision of the Institute}
To be a center of excellence in technical education and research, producing globally competent professionals who contribute to the advancement of technology and society.

\vspace{1cm}

\section*{Mission of the Institute}
\begin{itemize}
    \item To provide quality technical education through innovative teaching-learning processes.
    \item To promote research and development activities in emerging areas of technology.
    \item To foster industry-academia collaboration for practical exposure and skill development.
    \item To inculcate ethical values, leadership qualities, and social responsibility among students.
    \item To create an environment conducive to lifelong learning and professional growth.
\end{itemize}

\vspace{1cm}

\section*{Vision of the Department}
To emerge as a leading department in Computer Science and Engineering with specialization in Artificial Intelligence and Machine Learning, producing skilled professionals capable of addressing real-world challenges through innovative AI solutions.

\vspace{1cm}

\section*{Mission of the Department}
\begin{itemize}
    \item To provide comprehensive education in computer science fundamentals and advanced AI/ML technologies.
    \item To encourage research and innovation in artificial intelligence, machine learning, and related domains.
    \item To develop industry-ready professionals through hands-on projects and internships.
    \item To promote ethical AI practices and responsible technology development.
    \item To foster entrepreneurship and innovation in AI-driven solutions.
\end{itemize}

\newpage
